---
title: "Conceptual Figures"
author: "Joe Celebrezze"
date: "2023-09-15"
output: html_document
---
# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
here = here::here
library(tidyverse)
library(ggsci) #for scale_color_lancet
library(Ternary)

flamm.df <- read.csv(here('data', 'processed-data', 'main_dataset.csv'))
```

```{r}
flamm.df %>% 
  dplyr::filter(ignition != 0) %>% 
  drop_na(temp_change) %>% 
  dplyr::group_by(species) %>% 
  dplyr::select(species, fh, fd, temp_change, heat_flux_change, prop_ig, ignition) %>% 
  dplyr::summarise(fh = mean(fh), fd = mean(fd), temp_change = mean(temp_change),
            hfc = mean(heat_flux_change), prop_ig = mean(prop_ig))

flamm.df %>% 
  dplyr::group_by(species, ignition) %>% 
  tally()
```


# Survey-Traits-Flammability
First, to present idea to KF and MAM, going to just use subjective numbers (except for prop. ig.)
Ignitability
```{r}
prop_ig_df <- flamm.df %>%
  group_by(species) %>%
  summarise(prop_ig = mean(prop_ig)) %>% 
  mutate(prop_ig = prop_ig*2.25) # so maximum = 1
```

Dataframe
```{r}
spp <- c('ARTCAL', 'ARCDEN', 'CEAGRI', 'ERIKAR', 'HETARB', 'IRIDOU', 'MALLAU', 'SALAPI', 'SALLEU')
# flam.
comb <- c(0.3, 0.5, 0.7, 0.1, 1, 0, 0.9, 0.6, 0.5)
ign <- c(0.321, 0, 0.375, 0.09, 0, 0, 0.112, 1, 0.844)
# survey
drought <- c(1, 1, 1, 0.5, 1, 0, 1, 1, 1)
others <- c(0.5, 0.3, 0.6, 0.3, 0.5, 0.2, 0.5, 0.8, 0.8)
# traits
morph <- c(1, 0.3, 0.9, 0.9, 0.4, 0, 0.75, 0.6, 0.6)
hyd <- c(0.5, 0.2, 0.8, 0, 0.8, 0, 0.7, 0.6, 1)
# colors
color <- c('#ED0000B2', '#00468BB2', '#42B540B2', '#0099B4B2', '#925E9FB2', '#FDAF91B2', '#AD002AB2', '#ADB6B6B2', '#1B1919B2')

triangle.df <- data.frame(spp, comb, ign, drought, others, morph, hyd, color) %>% 
  mutate(flam = sqrt(comb^2 + ign^2)/1.1661904 + 0.1) %>% 
  mutate(survey = sqrt(drought^2 + others^2)/1.2806248 + 0.1) %>% 
  mutate(traits = sqrt(morph^2 + hyd^2)/1.2041595 + 0.1)
```

Plots
```{r}
par(mfrow = c(3, 3), mar = c(0.01, 0.01, 1, 0.01))
for(i in 1:nrow(triangle.df)){
TernaryPlot("Flammability", "Desirable Traits (Survey)", "Flam-Related Traits",
            grid.lines = 0,
            grid.minor.lines = 0,
            col = triangle.df$color[i])

title(triangle.df[i,1], cex.main = 1.5)

# Add an arrow
TernaryArrows(c(50, 50, 50), c(triangle.df[i,9], 0.1, 0.1), length = 0.14*triangle.df[i,9], lwd = 3*triangle.df[i,9], col = "black")
TernaryArrows(c(50, 50, 50), c(0.1, triangle.df[i,10], 0.1), length = 0.14*triangle.df[i,10], lwd = 3*triangle.df[i,10], col = "black")
TernaryArrows(c(50, 50, 50), c(0.1, 0.1, triangle.df[i,11]), length = 0.14*triangle.df[i,11], lwd = 3*triangle.df[i,11], col = "black")

# Color background
#cols <- TernaryPointValues(rgb, resolution = 100)
#ColourTernary(cols, spectrum = NULL)
}
```

# Combustibility-Ignitability-Survey/Traits
Dataframe (making minimum = 0.1, maximum = 1.1 forr combustibility, ignitability)
```{r}
triangle2.df <- triangle.df %>% 
  mutate(comb = comb + 0.1) %>% 
  mutate(ign = ign + 0.1)
```

Survey Plot:
```{r}
par(mfrow = c(3, 3), mar = c(0.01, 0.01, 1, 0.01))
for(i in 1:nrow(triangle2.df)){
TernaryPlot("Combustibility", "Desirable Traits (Survey)", "Ignitability",
            grid.lines = 0,
            grid.minor.lines = 0,
            col = triangle2.df$color[i])

title(triangle2.df[i,1], cex.main = 1.5)

# Add an arrow
TernaryArrows(c(50, 50, 50), c(triangle2.df[i,2], 0.1, 0.1), length = 0.14*triangle2.df[i,2], lwd = 3*triangle2.df[i,2], col = "black")
TernaryArrows(c(50, 50, 50), c(0.1, triangle2.df[i,9], 0.1), length = 0.14*triangle2.df[i,10],lwd = 3*triangle2.df[i,10], col = "black")
TernaryArrows(c(50, 50, 50), c(0.1, 0.1, triangle2.df[i,3]), length = 0.14*triangle2.df[i,3], lwd = 3*triangle2.df[i,3], col = "black")}
```

Traits Plot:
```{r}
par(mfrow = c(3, 3), mar = c(0.01, 0.01, 1, 0.01))
for(i in 1:nrow(triangle2.df)){
TernaryPlot("Combustibility", "Flam-Related Traits", "Ignitability",
            grid.lines = 0,
            grid.minor.lines = 0,
            col = triangle2.df$color[i])

title(triangle2.df[i,1], cex.main = 1.5)

# Add an arrow
TernaryArrows(c(50, 50, 50), c(triangle2.df[i,2], 0.1, 0.1), length = 0.14*triangle2.df[i,2], lwd = 3*triangle2.df[i,2], col = "black")
TernaryArrows(c(50, 50, 50), c(0.1, triangle2.df[i,10], 0.1), length = 0.14*triangle2.df[i,11], lwd = 3*triangle2.df[i,11], col = "black")
TernaryArrows(c(50, 50, 50), c(0.1, 0.1, triangle2.df[i,3]), length = 0.14*triangle2.df[i,3], lwd = 3*triangle2.df[i,3], col = "black")}
```

# Combustibility vs. Ignitability
```{r}
ggplot(data = triangle.df, aes(x = comb, y = ign, color = spp)) +
  geom_text(label = spp, size = 8, fontface = 'bold') +
  labs(x = 'Combustibility', y = 'Ignitablity') +
  theme_classic() +
  coord_equal() +
  scale_color_lancet() +
  theme(legend.position = 'none',
        axis.title = element_text(face = 'bold', size = 30),
        axis.text = element_blank(),
        axis.ticks = element_blank())
ggsave(here('figures', 'main_figures', 'conceptual.ig.vs.comb.png'), height = 16, width = 16)
```

# Fire Triangles
Creating indices for each of the three axes (ignitability, combustibility, sustainablity)
```{r}
flamm.means <- flamm.df %>% 
  drop_na(fh, fd, prop_ig, temp_change, heat_flux_change) %>% 
  summarise(fh_mean = mean(fh), fh_sd = sd(fh),
            fd_mean = mean(fd), fd_sd = sd(fd),
            prop_ig_mean = mean(prop_ig), prop_ig_sd = sd(prop_ig),
            temp_change_mean = mean(temp_change), temp_change_sd = sd(temp_change),
            heat_flux_change_mean = mean(heat_flux_change), heat_flux_change_sd = sd(heat_flux_change))
flamm.index <- flamm.df %>% 
  drop_na(fh, fd, prop_ig, temp_change, heat_flux_change) %>% 
  mutate(fh_scale = (fh - flamm.means$fh_mean[1])/flamm.means$fh_sd[1],
         heat_flux_change_scale = (heat_flux_change - flamm.means$heat_flux_change_mean[1])/flamm.means$heat_flux_change_sd[1],
         temp_change_scale = (temp_change - flamm.means$temp_change_mean[1])/flamm.means$temp_change_sd[1],
         fd_scale = (fd - flamm.means$fd_mean[1])/flamm.means$fd_sd[1],
         prop_ig_scale = (prop_ig - flamm.means$prop_ig_mean[1])/flamm.means$prop_ig_sd[1]) %>% 
  group_by(species) %>% 
  summarise(fh = mean(fh_scale), fd = mean(fd_scale), prop_ig = mean(prop_ig_scale),
            temp_change = mean(temp_change_scale), heat_flux_change = mean(heat_flux_change_scale)) %>% 
  mutate(ignitability = (prop_ig + 1.1154)/2.83, combustibility = (((fh + heat_flux_change + temp_change)/3) + 0.8667)/2.42, sustainability = (fd + 0.76)/1.606) %>% 
  dplyr::select(species, ignitability, combustibility, sustainability) %>% 
  ungroup()

# color
flamm.index$color <- c('#00468BB2', '#ED0000B2', '#42B540B2', '#0099B4B2', '#925E9FB2', '#FDAF91B2', '#AD002AB2', '#ADB6B6B2', '#1B1919B2')
```

Plotting
```{r}
par(mfrow = c(3, 3), mar = c(0.01, 0.01, 1.3, 0.01))
for(i in 1:nrow(flamm.index)){
TernaryPlot("Ignitability", "Combustibility", "Sustainabilty",
            grid.lines = 0,
            grid.minor.lines = 0,
            lab.cex = 2,
            col = flamm.index$color[i])

title(flamm.index[i,1], cex.main = 2)

# Add an arrow
TernaryArrows(c(50, 50, 50), c(flamm.index$ignitability[i], 0.03, 0.03), length = 0.2*flamm.index$ignitability[i], lwd = 5*flamm.index$ignitability[i], col = "black")
TernaryArrows(c(50, 50, 50), c(0.03, flamm.index$combustibility[i], 0.03), length = 0.2*flamm.index$combustibility[i], lwd = 5*flamm.index$combustibility[i], col = "black")
TernaryArrows(c(50, 50, 50), c(0.03, 0.03, flamm.index$sustainability[i]), length = 0.2*flamm.index$sustainability[i], lwd = 5*flamm.index$sustainability[i], col = "black")
}
```

## REMEF Version
```{r}
conceptual_flam_remef <- read.csv(here('data', 'processed-data', 'conceptual_flam_remef.csv'))

r.flamm.index <- conceptual_flam_remef %>% 
  drop_na(r_fh, r_fd, prop_ig, r_temp_change, r_heat_flux_change) %>% 
  group_by(species) %>% 
  summarise(fh = mean(r_fh), fd = mean(r_fd), prop_ig = mean(prop_ig),
            temp_change = mean(r_temp_change), heat_flux_change = mean(r_heat_flux_change)) %>% 
  mutate(ignitability = (prop_ig + 1.1154)/2.83, combustibility = (((fh + heat_flux_change + temp_change)/3) + 0.8667)/2.42, sustainability = (fd + 0.76)/1.606) %>% 
  dplyr::select(species, ignitability, combustibility, sustainability) %>% 
  ungroup()

# color
r.flamm.index$color <- c('#00468BB2', '#ED0000B2', '#42B540B2', '#0099B4B2', '#925E9FB2', '#AD002AB2', '#ADB6B6B2', '#1B1919B2')
```

Plotting
```{r}
par(mfrow = c(3, 3), mar = c(0.01, 0.01, 1.3, 0.01))
for(i in 1:nrow(r.flamm.index)){
TernaryPlot("Ignitability", "Combustibility", "Sustainabilty",
            grid.lines = 0,
            grid.minor.lines = 0,
            lab.cex = 2,
            col = r.flamm.index$color[i])

title(r.flamm.index[i,1], cex.main = 2)

# Add an arrow
TernaryArrows(c(50, 50, 50), c(r.flamm.index$ignitability[i], 0.1, 0.1), length = 0.3*r.flamm.index$ignitability[i], lwd = 5*r.flamm.index$ignitability[i], col = "black")
TernaryArrows(c(50, 50, 50), c(0.1, r.flamm.index$combustibility[i], 0.1), length = 0.2*r.flamm.index$combustibility[i], lwd = 5*r.flamm.index$combustibility[i], col = "black")
TernaryArrows(c(50, 50, 50), c(0.1, 0.1, r.flamm.index$sustainability[i]), length = 0.2*r.flamm.index$sustainability[i], lwd = 5*r.flamm.index$sustainability[i], col = "black")
}
```