---
title: "Data Wrangling: Plant Traits"
author: "Zoe Fung & Joe Celebrezze"
date: "2023-04-17"
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(janitor)
library(dplyr)
library(here)
```

# Initial Data Cleaning
## Leaf Data
```{r}
# load in data
leaf_data <- read_csv(here('data', 'raw-data', 'plant-traits', 'leaf_data_flamm.csv')) %>%
  clean_names() %>%
  slice_head(n = 467) %>%
  select(-x12)

# view column names and data types
glimpse(leaf_data)
```

## Leaf Area
```{r}
# load in data
leaf_area <- read_csv(here('data', 'raw-data', 'plant-traits', 'leaf_area_flamm.csv')) %>%
  clean_names() %>%
  slice_head(n = 314) %>%
  select(-c(x_1, x30, x31))

# view column names and data types
glimpse(leaf_area)
```

## Stem-Leaf Ratio
```{r}
# load in data
stem_leaf_ratio <- read_csv(here('data', 'raw-data', 'plant-traits', 'stemleaf_massratio_flamm.csv')) %>%
  clean_names() %>%
  slice_head(n = 92) %>%
  select(-c(x10, x11, x12))

# view column names and data types
glimpse(stem_leaf_ratio)
```

## Stem Data
```{r}
# load in data
stem_data <- read_csv(here('data', 'raw-data', 'plant-traits', 'stem_data_flamm.csv')) %>%
  clean_names() %>%
  slice_head(n = 371) 

# view column names and data types
glimpse(stem_data)
```

# Calculating LFM
First, getting cup weights data
```{r}
cup_weights_data <- read_csv(here('data', 'raw-data', 'plant-traits', 'cup_weights.csv')) %>% 
  slice_head(n = 82)  %>% 
  rename(cup_mass = mass)

leaf_data <- merge(leaf_data, cup_weights_data, by = 'cup_number')
stem_data <- merge(stem_data, cup_weights_data, by = 'cup_number')
```

Subtracting cup weight from fresh and dry weights and then calculating stem- and leaf-specific LFM
```{r}
leaf_data <- leaf_data %>% 
  mutate(leaf_wet_mass_g = ifelse(leaf_wet_mass_g > 1000, leaf_wet_mass_g/1000, leaf_wet_mass_g)) %>% 
  mutate(leaf_wet_mass = leaf_wet_mass_g - cup_mass) %>% 
  mutate(leaf_dry_mass = leaf_dry_mass - cup_mass) %>% 
  select(-leaf_wet_mass_g) %>% 
  mutate(leaf_lfm = 100*(leaf_wet_mass - leaf_dry_mass)/leaf_dry_mass)

stem_data <- stem_data %>% 
  mutate(stem_wet_mass = stem_wet_mass - cup_mass) %>% 
  mutate(stem_dry_mass = stem_dry_mass - cup_mass) %>% 
  mutate(stem_lfm = 100*(stem_wet_mass - stem_dry_mass)/stem_dry_mass)
```

# Data Wrangling
Adding some variables
```{r}
stem_data <- stem_data %>% 
  mutate(stem_width = (stem_width1 + stem_width2)/2) #average of two stem width measurements taken

leaf_area <- leaf_area %>% 
  mutate(plant = case_when(
    plant_number == 1 ~ 'P1',
    plant_number == 2 ~ 'P2',
    plant_number == 3 ~ 'P3',
    plant_number == 4 ~ 'P4',
    plant_number == 5 ~ 'P5',
    plant_number == 6 ~ 'P6',
    plant_number == 7 ~ 'P7',
    plant_number == 8 ~ 'P8',
    plant_number == 9 ~ 'P9',
    plant_number == 10 ~ 'P10'
  ))
```

Averaging traits of interest so that there is one value per plant
```{r}
stem_data_avgs <- stem_data %>% 
  group_by(species, plant) %>% 
  summarise(stem_lfm = mean(stem_lfm), stem_width = mean(stem_width), stem_wet_mass = mean(stem_wet_mass), stem_dry_mass = mean(stem_dry_mass))

leaf_data_avgs <- leaf_data %>% 
  group_by(species, plant) %>% 
  summarise(leaf_lfm = mean(leaf_lfm), thickness = mean(thickness), leaf_wet_mass = mean(leaf_wet_mass), leaf_dry_mass = mean(leaf_dry_mass))

leaf_area_avgs <- leaf_area %>% 
  group_by(species, plant) %>% 
  summarise(area_cm2 = mean(area_cm2), round = mean(round), solidity = mean(solidity))
```

Combining datasets so that they can be easily integrated into the flamm data

- First leaf_data and stem_data
```{r}
## Adding IRIDOU to stem_data w/ NAs
species <- c(rep('IRIDOU', 9))
plant <- c('P1', 'P2', 'P3', 'P4', 'P5', 'P6', 'P7', 'P8', 'P9')
na <- c(rep(999999, 9)) # so that IRIDOU data does not get lost in the merge
stem_data_avgs_iridou <- data.frame(species, plant, na, na, na, na)
colnames(stem_data_avgs_iridou) <- colnames(stem_data_avgs)
stem_data_avgs <- rbind(stem_data_avgs, stem_data_avgs_iridou)
## Combining
leaf_stem_df <- merge(leaf_data_avgs, stem_data_avgs, by = c('species', 'plant'))
```

- Calculating total LFM
```{r}
leaf_stem_df <- leaf_stem_df %>% 
  mutate(lfm = 100*((stem_wet_mass + leaf_wet_mass) - (stem_dry_mass + leaf_dry_mass))/(stem_dry_mass + leaf_dry_mass))

leaf_stem_df %>% 
  filter(species != 'IRIDOU') %>% 
  filter(lfm > 0  & lfm < 500) %>% 
  ggplot(aes(x = lfm, y = stem_lfm)) +
    labs(x = 'LFM', y = 'LFM (stem)') +
    geom_point() +
    geom_smooth(method = 'lm', color = 'black') +
    geom_abline(linetype = 2) +
    theme_bw()

leaf_stem_df %>% 
  filter(species != 'IRIDOU') %>% 
  filter(lfm > 0  & lfm < 500) %>% 
  filter(leaf_lfm > 0  & leaf_lfm < 900) %>% 
  ggplot(aes(x = lfm, y = leaf_lfm)) +
    labs(x = 'LFM', y = 'LFM (leaf)') +
    geom_point() +
    geom_smooth(method = 'lm', color = 'black') +
    geom_abline(linetype = 2) +
    theme_bw()

leaf_stem_df %>% 
  filter(species != 'IRIDOU') %>% 
  filter(leaf_lfm > 0  & leaf_lfm < 900) %>% 
  filter(stem_lfm > 0  & stem_lfm < 600) %>% 
  ggplot(aes(x = stem_lfm, y = leaf_lfm)) +
    labs(x = 'LFM (stem)', y = 'LFM (leaf)') +
    geom_point() +
    geom_smooth(method = 'lm', color = 'black') +
    geom_abline(linetype = 2) +
    theme_bw()
```

- Now, stem_leaf_ratio
note: for now, not going to add or takeaway any data prior to merging, so we will lose some of the data from the above dataset. Hopefully, Zoe, Rob and I will figure this out by looking into where the missing trait data could be.
```{r}
stem_leaf_ratio_clean <- stem_leaf_ratio %>% 
  select(species, plant, date, total_branch_mass, total_leaf_mass) %>% 
  mutate(stem_mass_ratio = total_branch_mass/(total_branch_mass + total_leaf_mass)) %>% 
  mutate(leaf_mass_ratio = total_leaf_mass/(total_branch_mass + total_leaf_mass))

leaf_stem_v2_df <- merge(stem_leaf_ratio_clean, leaf_stem_df, by = c('species', 'plant'))
```

- Missing a bunch of leaf_area data, so not going to merge that quite yet. I want to calculate LMA and SLA as I assume we'll want to look at those, but I don't know if I can. Is leaf_wet_mass only for five leaves we took area on?
`