---
title: "Data Wrangling: Plant Traits"
author: "Zoe Fung & Joe Celebrezze"
date: "2023-04-17"
output: html_document
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(janitor)
library(dplyr)
library(here)
```

# Initial Data Cleaning
## Leaf Data
```{r}
# load in data
leaf_data <- read_csv(here('data', 'raw-data', 'plant-traits', 'leaf_data_flamm.csv')) %>%
  clean_names()

# view column names and data types
glimpse(leaf_data)
```

## Leaf Area
```{r}
# load in data
leaf_area <- read_csv(here('data', 'raw-data', 'plant-traits', 'leaf_area_flamm.csv')) %>% 
  clean_names() %>% 
  select(-x_1)

# view column names and data types
glimpse(leaf_area)
```

## Stem-Leaf Ratio
```{r}
# load in data
stem_leaf_ratio <- read_csv(here('data', 'raw-data', 'plant-traits', 'stemleaf_massratio_flamm.csv')) %>%
  clean_names()

# view column names and data types
glimpse(stem_leaf_ratio)
```

## Stem Data
```{r}
# load in data
stem_data <- read_csv(here('data', 'raw-data', 'plant-traits', 'stem_data_flamm.csv')) %>%
  clean_names()

# view column names and data types
glimpse(stem_data)
```

# Calculating LFM
First, getting cup weights data
```{r}
cup_weights_data <- read_csv(here('data', 'raw-data', 'plant-traits', 'cup_weights.csv')) %>% 
  slice_head(n = 82)  %>% 
  rename(cup_mass = mass)

leaf_data <- merge(leaf_data, cup_weights_data, by = 'cup_number')
stem_data <- merge(stem_data, cup_weights_data, by = 'cup_number')
```

Subtracting cup weight from fresh and dry weights and then calculating stem- and leaf-specific LFM
```{r}
leaf_data <- leaf_data %>% 
  mutate(leaf_wet_mass_g = ifelse(leaf_wet_mass_g > 1000, leaf_wet_mass_g/1000, leaf_wet_mass_g)) %>% 
  mutate(leaf_wet_mass = leaf_wet_mass_g - cup_mass) %>% 
  mutate(leaf_dry_mass = leaf_dry_mass - cup_mass) %>% 
  select(-leaf_wet_mass_g) %>% 
  mutate(leaf_lfm = 100*(leaf_wet_mass - leaf_dry_mass)/leaf_dry_mass)

stem_data <- stem_data %>% 
  mutate(stem_wet_mass = stem_wet_mass - cup_mass) %>% 
  mutate(stem_dry_mass = stem_dry_mass - cup_mass) %>% 
  mutate(stem_lfm = 100*(stem_wet_mass - stem_dry_mass)/stem_dry_mass)
```

# Data Wrangling

```{r}
stem_data <- stem_data %>% 
  mutate(stem_width = (stem_width1 + stem_width2)/2) #average of two stem width measurements taken

leaf_area <- leaf_area %>% 
  mutate(plant = case_when(
    plant_number == 1 ~ 'P1',
    plant_number == 2 ~ 'P2',
    plant_number == 3 ~ 'P3',
    plant_number == 4 ~ 'P4',
    plant_number == 5 ~ 'P5',
    plant_number == 6 ~ 'P6',
    plant_number == 7 ~ 'P7',
    plant_number == 8 ~ 'P8',
    plant_number == 9 ~ 'P9',
    plant_number == 10 ~ 'P10',
    plant_number == 11 ~ 'P11',
    plant_number == 12 ~ 'P12'
  ))
```

## Averaging Traits
Averaging traits of interest so that there is one value per plant
```{r}
stem_data_avgs <- stem_data %>% 
  group_by(species, plant) %>% 
  summarise(stem_lfm = mean(stem_lfm), stem_width = mean(stem_width), stem_wet_mass = mean(stem_wet_mass), stem_dry_mass = mean(stem_dry_mass))

leaf_data_avgs <- leaf_data %>% 
  group_by(species, plant) %>% 
  summarise(leaf_lfm = mean(leaf_lfm), thickness = mean(thickness), leaf_wet_mass = mean(leaf_wet_mass), leaf_dry_mass = mean(leaf_dry_mass))

leaf_area_avgs <- leaf_area %>% 
  group_by(species, plant) %>% 
  summarise(leaf_area = mean(area_cm2))
```

## Combining Traits Data
Combining datasets so that they can be easily integrated into the flamm data

- First leaf_data and stem_data
```{r}
## Adding IRIDOU to stem_data w/ NAs
species <- c(rep('IRIDOU', 9))
plant <- c('P1', 'P2', 'P3', 'P4', 'P5', 'P6', 'P7', 'P8', 'P9')
na <- c(rep(999999, 9)) # so that IRIDOU data does not get lost in the merge
stem_data_avgs_iridou <- data.frame(species, plant, na, na, na, na)
colnames(stem_data_avgs_iridou) <- colnames(stem_data_avgs)
stem_data_avgs <- rbind(stem_data_avgs, stem_data_avgs_iridou)
## Combining
leaf_stem_df <- merge(leaf_data_avgs, stem_data_avgs, by = c('species', 'plant'))
```

### Calculating total LFM
```{r}
leaf_stem_df <- leaf_stem_df %>% 
  mutate(lfm = 100*((stem_wet_mass + leaf_wet_mass) - (stem_dry_mass + leaf_dry_mass))/(stem_dry_mass + leaf_dry_mass))

leaf_stem_df %>% 
  filter(species != 'IRIDOU') %>% 
  filter(lfm > 0  & lfm < 500) %>% 
  ggplot(aes(x = lfm, y = stem_lfm)) +
    labs(x = 'LFM', y = 'LFM (stem)') +
    geom_point() +
    geom_smooth(method = 'lm', color = 'black') +
    geom_abline(linetype = 2) +
    theme_bw()

leaf_stem_df %>% 
  filter(species != 'IRIDOU') %>% 
  filter(lfm > 0  & lfm < 500) %>% 
  filter(leaf_lfm > 0  & leaf_lfm < 900) %>% 
  ggplot(aes(x = lfm, y = leaf_lfm)) +
    labs(x = 'LFM', y = 'LFM (leaf)') +
    geom_point() +
    geom_smooth(method = 'lm', color = 'black') +
    geom_abline(linetype = 2) +
    theme_bw()

leaf_stem_df %>% 
  filter(species != 'IRIDOU') %>% 
  filter(leaf_lfm > 0  & leaf_lfm < 900) %>% 
  filter(stem_lfm > 0  & stem_lfm < 600) %>% 
  ggplot(aes(x = stem_lfm, y = leaf_lfm)) +
    labs(x = 'LFM (stem)', y = 'LFM (leaf)') +
    geom_point() +
    geom_smooth(method = 'lm', color = 'black') +
    geom_abline(linetype = 2) +
    theme_bw()
```

- Now, stem_leaf_ratio
note: for now, not going to add or takeaway any data prior to merging, so we will lose some of the data from the above dataset. Hopefully, Zoe, Rob and I will figure this out by looking into where the missing trait data could be.
```{r}
stem_leaf_ratio_clean <- stem_leaf_ratio %>% 
  select(species, plant, date, total_branch_mass, total_leaf_mass) %>% 
  mutate(stem_mass_ratio = total_branch_mass/(total_branch_mass + total_leaf_mass)) %>% 
  mutate(leaf_mass_ratio = total_leaf_mass/(total_branch_mass + total_leaf_mass))

leaf_stem_v2_df <- merge(stem_leaf_ratio_clean, leaf_stem_df, by = c('species', 'plant'))
```

- Adding leaf area
```{r}
# Adding NA values for HETARB P10-12
species <- c(rep('HETARB', 3))
plant <- c('P10', 'P11', 'P12')
na <- c(rep(999999, 3)) # so that IRIDOU data does not get lost in the merge
leaf_area_avgs_hetarb <- data.frame(species, plant, na)
colnames(leaf_area_avgs_hetarb) <- colnames(leaf_area_avgs)
leaf_area_avgs <- rbind(leaf_area_avgs, leaf_area_avgs_hetarb)

main_traits_df <- merge(leaf_stem_v2_df, leaf_area_avgs, by = c('species', 'plant'))
```

- Dealing with NAs from above (HETARB, IRIDOU)
```{r}
main_traits_df <- main_traits_df %>% 
  mutate(stem_lfm = ifelse(stem_lfm > 99999, NA, stem_lfm)) %>% 
  mutate(stem_width = ifelse(stem_width > 99999, NA, stem_width)) %>%
  mutate(stem_wet_mass = ifelse(stem_wet_mass > 99999, NA, stem_wet_mass)) %>% 
  mutate(stem_dry_mass = ifelse(stem_dry_mass > 99999, NA, stem_dry_mass)) %>% 
  mutate(lfm = ifelse(stem_lfm > 99999, NA, lfm)) %>% 
  mutate(leaf_area = ifelse(leaf_area > 99999, NA, leaf_area))
```

## Calculating LMA, SLA
```{r}
main_traits_df <- main_traits_df %>% 
  mutate(LMA = leaf_wet_mass/leaf_area) %>% 
  mutate(SLA = leaf_area/leaf_wet_mass)
```

## Small Adjustments
```{r}
main_traits_df <- main_traits_df %>% 
  mutate(plant = case_when(
    plant == 'P1' ~ 1,
    plant == 'P2' ~ 2,
    plant == 'P3' ~ 3,
    plant == 'P4' ~ 4,
    plant == 'P5' ~ 5,
    plant == 'P6' ~ 6,
    plant == 'P7' ~ 7,
    plant == 'P8' ~ 8,
    plant == 'P9' ~ 9,
    plant == 'P10' ~ 10,
    plant == 'P11' ~ 11,
    plant == 'P12' ~ 12
  )) %>% 
  mutate(species = ifelse(species == 'EUGL', 'EUCGLO', species))
```

# Writing csv
```{r}
write.csv(main_traits_df, here('data', 'processed-data', 'plant_traits_data.csv'))
```

# Combining with Flam. Dataset
```{r}
flam.df <- read.csv(here('data', 'processed-data', 'clean_flamm_data.csv')) %>% 
  select(-date)

main_df <- merge(main_traits_df, flam.df, by = c('species', 'plant'))

write.csv(main_df, here('data', 'processed-data', 'main_plant_flammability.csv'))
```

